---
title: "Discrete choice models"
author: "C Markwalter"
date: "2023-10-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries and data
```{r}
library(biostat3)
library(tidyverse)
library(mlogit)
library(broom)
library(ggeffects)
library(ggpubr)

model_dat_long <- read_csv("analysis/clean_data/data/model_dat_long.csv")
```

# Function for shaping data

Note: the scalar for determining sporozoite load comes from the following scalars:

Going to convert our ng/ul concentrations to genomes in order to get more relevant units on our spz densities. This conversion is based on the following:

1 pg ~ 978 Mbp
Pf genome 22.8 Mbp
1 ul template -> units are ng

So 1 ng = 1000 pg * 978 Mbp/pg * 1 genome/22.8 Mbp = 42894.74 genomes

```{r}

get_modeldat <- function(df, hu_var){
  
  # reshaping multisource
  
  # first duplicate all the 0's
  multi_no_match <- df %>%
        select(sample_id_mosquito, match, age_cat, gender, slept_net, pf_pcr_infection_status, moz_head_pf_density, moz_head_pf_status, species, missing_data, multisource, locus_count) %>%
      filter(species %in% c("An. funestus", "An. gambiae SS")) %>% 
      mutate(moz_head_pf_density = case_when(is.na(moz_head_pf_density) & moz_head_pf_status == "negative" ~ 0, !is.na(moz_head_pf_density) ~  moz_head_pf_density * 42894.74),
             slept_net = factor(slept_net, levels = rev(c("Yes", "No")))) %>%
      drop_na() %>%
      group_by(sample_id_mosquito) %>%
      filter(sum(match == "yes") > 1) %>%
        ungroup() %>%
        filter(match == "No") %>%
        mutate(sample_id_mosquito = paste0(sample_id_mosquito, "_1")) %>%
        bind_rows(df %>%
      select(sample_id_mosquito, match, age_cat, gender, slept_net, pf_pcr_infection_status, moz_head_pf_density, moz_head_pf_status, species, missing_data, multisource, locus_count) %>%
      filter(species %in% c("An. funestus", "An. gambiae SS")) %>% 
      mutate(moz_head_pf_density = case_when(is.na(moz_head_pf_density) & moz_head_pf_status == "negative" ~ 0, !is.na(moz_head_pf_density) ~  moz_head_pf_density * 42894.74),
             slept_net = factor(slept_net, levels = rev(c("Yes", "No")))) %>%
      drop_na() %>%
      group_by(sample_id_mosquito) %>%
      filter(sum(match == "yes") > 1) %>%
        ungroup() %>%
        filter(match == "No") %>%
        mutate(sample_id_mosquito = paste0(sample_id_mosquito, "_2")))
  
  # assign matches to new unique mozzie ids and add the appropriate 0's
  multi <- df %>%
      select(sample_id_mosquito, match, age_cat, gender, slept_net, pf_pcr_infection_status, moz_head_pf_density, moz_head_pf_status, species, missing_data, multisource, locus_count) %>%
      filter(species %in% c("An. funestus", "An. gambiae SS")) %>% 
      mutate(moz_head_pf_density = case_when(is.na(moz_head_pf_density) & moz_head_pf_status == "negative" ~ 0, !is.na(moz_head_pf_density) ~  moz_head_pf_density * 42894.74),
             slept_net = factor(slept_net, levels = rev(c("Yes", "No")))) %>%
      drop_na() %>%
      group_by(sample_id_mosquito) %>%
      filter(sum(match == "yes") > 1) %>%
      mutate(hh_index = row_number()) %>%
        ungroup() %>%
        filter(match == "yes") %>%
        group_by(sample_id_mosquito) %>%
        mutate(sample_id_mosquito = case_when(hh_index == min(hh_index) ~ paste0(sample_id_mosquito, "_1"),
                                                 hh_index == max(hh_index) ~ paste0(sample_id_mosquito, "_2"))) %>%
        ungroup() %>%
        bind_rows(df %>%
      select(sample_id_mosquito, match, age_cat, gender, slept_net, pf_pcr_infection_status, moz_head_pf_density, moz_head_pf_status, species, missing_data, multisource, locus_count) %>%
      filter(species %in% c("An. funestus", "An. gambiae SS")) %>% 
      mutate(moz_head_pf_density = case_when(is.na(moz_head_pf_density) & moz_head_pf_status == "negative" ~ 0, !is.na(moz_head_pf_density) ~  moz_head_pf_density * 42894.74),
             slept_net = factor(slept_net, levels = rev(c("Yes", "No")))) %>%
      drop_na() %>%
      group_by(sample_id_mosquito) %>%
      filter(sum(match == "yes") > 1) %>%
      mutate(hh_index = row_number()) %>%
        ungroup() %>%
        filter(match == "yes") %>%
        group_by(sample_id_mosquito) %>%
        mutate(sample_id_mosquito = case_when(hh_index == min(hh_index) ~ paste0(sample_id_mosquito, "_2"),
                                                 hh_index == max(hh_index) ~ paste0(sample_id_mosquito, "_1")),
               match = "No") %>%
        ungroup()) %>%
        bind_rows(multi_no_match)
  
  modeldat <- df %>%
      select(sample_id_mosquito, match, age_cat, gender, slept_net, pf_pcr_infection_status, moz_head_pf_density, moz_head_pf_status, species, missing_data, multisource, locus_count) %>%
      filter(species %in% c("An. funestus", "An. gambiae SS")) %>% 
      mutate(moz_head_pf_density = case_when(is.na(moz_head_pf_density) & moz_head_pf_status == "negative" ~ 0, !is.na(moz_head_pf_density) ~  moz_head_pf_density * 42894.74),
             slept_net = factor(slept_net, levels = rev(c("Yes", "No")))) %>%
      drop_na() %>%
      group_by(sample_id_mosquito) %>%
      filter(sum(match == "yes") == 1) %>%
      ungroup() %>%
      bind_rows(multi) %>%
      mutate(index_mosquito = as.numeric(factor(sample_id_mosquito)),
             choice = as.numeric(match == "yes"),
             species = factor(species, levels = c("An. gambiae SS", "An. funestus"))) %>% 
      group_by(index_mosquito, {{hu_var}}) %>%
      mutate(n = n(),
             choice = case_when(sum(choice) > 0 ~ 1,
                                TRUE ~ 0)) %>%
      select(sample_id_mosquito, {{hu_var}}, moz_head_pf_density, moz_head_pf_status, species, multisource, missing_data,multisource, locus_count, index_mosquito, choice, n) %>%
      distinct() %>%
      group_by(index_mosquito) %>%
      mutate(weight = n/sum(n)) %>%
      ungroup()

  return(modeldat)
}


```

# Get data

```{r}
gender <- get_modeldat(df = model_dat_long, hu_var = gender) 

age <- get_modeldat(df = model_dat_long, hu_var = age_cat) %>%
  mutate(age_cat = factor(age_cat, levels = c(">15", "<5", "5-15")))

hu_inf <- get_modeldat(df = model_dat_long, hu_var = pf_pcr_infection_status)

net <- get_modeldat(df = model_dat_long, hu_var = slept_net)
```


# Cat Spz: all mozzie covariates

## Hu infection
```{r}
m_inf_all <- mlogit(choice ~ 1 | multisource + moz_head_pf_status + species, chid.var = "index_mosquito", alt.var = 'pf_pcr_infection_status', weights = weight, data = hu_inf)

summary(m_inf_all)

exp(coef(m_inf_all))
```


## Gender

```{r}
m_gender_all <- mlogit(choice ~ 1 | multisource + moz_head_pf_status + species, chid.var = "index_mosquito", alt.var = 'gender', weights = weight, data = gender)

summary(m_gender_all)

exp(coef(m_gender_all))
```

## Net

```{r}
m_net_all <- mlogit(choice ~ 1 | multisource + moz_head_pf_status + species, chid.var = "index_mosquito", alt.var = 'slept_net', weights = weight, data = net)

summary(m_net_all)

exp(coef(m_net_all))
```

## Age

```{r}
m_age_all <- mlogit(choice ~ 1 | multisource + moz_head_pf_status + species, chid.var = "index_mosquito", alt.var = 'age_cat', weights = weight, data = age)

summary(m_age_all)

exp(coef(m_age_all))
```

## Put it together

```{r}
m_out_all <- bind_rows(tidy(m_gender_all, conf.int = TRUE),
                   tidy(m_age_all, conf.int = TRUE),
                   tidy(m_inf_all, conf.int = TRUE),
                   tidy(m_net_all, conf.int = TRUE)) %>%
  filter(!grepl('Intercept', term)) %>%
  mutate(estimate = exp(estimate),
         conf.low = exp(conf.low),
         conf.high = exp(conf.high),
         hu_char = gsub('.*:', '', term), 
         age = case_when(hu_char == "<5" ~ "< 5y",
                         hu_char == "5-15" ~ "5 - 15y"),
         hu_char = case_when(hu_char == 'Yes' ~ 'Net\nvs. no net (ref)',
                             hu_char == "positive" ~ 'Pf infection vs.\nNo Pf (ref)',
                             hu_char == "<5" ~ "Age vs.\n> 15y (ref)",
                             hu_char == "5-15" ~ "Age vs.\n> 15y (ref)",
                             hu_char == "Male" ~ "Male vs.\nFemale (ref)"),
         hu_char = factor(hu_char, levels = c("Age vs.\n> 15y (ref)", "Male vs.\nFemale (ref)", "Net\nvs. no net (ref)", 
"Pf infection vs.\nNo Pf (ref)")),
         moz_char = gsub(':.*', '', term),
         moz_char = case_when(moz_char == "speciesAn. funestus" ~ "Mosquito species:\nAn. funestus vs.\nAn. gambiae SS (ref)",
                              moz_char == "multisourceYes" ~ "Multisource bloodmeal:\n Yes vs. No (ref)",
                              moz_char == "moz_head_pf_statuspositive" ~ "Sporozoites in head-thorax:\npositive vs negative (ref)",
                              moz_char == "moz_head_pf_density" ~ "Sporozoites in\nhead-thorax"),
moz_char = factor(moz_char, levels = c("Mosquito species:\nAn. funestus vs.\nAn. gambiae SS (ref)", 
"Multisource bloodmeal:\n Yes vs. No (ref)", 
"Sporozoites in head-thorax:\npositive vs negative (ref)", "Sporozoites in\nhead-thorax"))) %>%
  select(-term)

m_out_all
```

```{r}
moz_plot <- m_out_all %>%
  mutate(sig = case_when(p.value < 0.05 ~ 1,
                         TRUE ~ 0)) %>%
  ggplot() +
  facet_grid(.~fct_rev(hu_char), scales = "free") +
  geom_vline(xintercept = 1, linetype = 'dashed') +
  geom_pointrange(aes(x = estimate, xmin = conf.low, xmax = conf.high, y = fct_rev(moz_char), color = fct_rev(age)), position = position_dodge2(width = 0.5)) +
  scale_color_manual(values = c("darkred", "navy"), breaks = c("< 5y", "5 - 15y"), na.value = "black")+
  labs(x = "Relative risk ratio", y = "Mosquito characteristic", title = "Human characteristic", color = "Age") +
  scale_x_continuous(trans = "log", breaks =  scales::breaks_extended(n = 4)) +
  theme_minimal() +
  theme(strip.text.y = element_blank(), panel.border = element_rect(color = "black", fill = NA), axis.text.y = element_text(color = "black"), plot.title = element_text(hjust = 0.5, size = 11))
```

# Continuous Spz: all mozzie covariates

## Hu infection
```{r}
m_inf_all_spz <- mlogit(choice ~ 1 | multisource + moz_head_pf_density + species, chid.var = "index_mosquito", alt.var = 'pf_pcr_infection_status', weights = weight, data = hu_inf)

summary(m_inf_all_spz)

exp(coef(m_inf_all_spz))
```

#### Per 100 spz

```{r}
m_inf_all_100spz <- mlogit(choice ~ 1 | multisource + moz_head_pf_density + species, chid.var = "index_mosquito", alt.var = 'pf_pcr_infection_status', weights = weight, data = hu_inf %>% mutate(moz_head_pf_density = moz_head_pf_density/100))

summary(m_inf_all_100spz)

exp(coef(m_inf_all_100spz))

tidy(m_inf_all_100spz, conf.int = TRUE) %>%
  mutate(estimate = exp(estimate),
         conf.low = exp(conf.low),
         conf.high = exp(conf.high))
```



## Gender

```{r}
m_gender_all_spz <- mlogit(choice ~ 1 | multisource + moz_head_pf_density + species, chid.var = "index_mosquito", alt.var = 'gender', weights = weight, data = gender)

summary(m_gender_all_spz)

exp(coef(m_gender_all_spz))
```

## Net

```{r}
m_net_all_spz <- mlogit(choice ~ 1 | multisource + moz_head_pf_density + species, chid.var = "index_mosquito", alt.var = 'slept_net', weights = weight, data = net)

summary(m_net_all_spz)

exp(coef(m_net_all_spz))
```

## Age

```{r}
m_age_all_spz <- mlogit(choice ~ 1 | multisource + moz_head_pf_density + species, chid.var = "index_mosquito", alt.var = 'age_cat', weights = weight, data = age)

summary(m_age_all_spz)

exp(coef(m_age_all_spz))
```


## Put it together

```{r}
m_out_all_spz <- bind_rows(tidy(m_gender_all_spz, conf.int = TRUE),
                   tidy(m_age_all_spz, conf.int = TRUE),
                   tidy(m_inf_all_spz, conf.int = TRUE),
                   tidy(m_net_all_spz, conf.int = TRUE)) %>%
  filter(!grepl('Intercept', term)) %>%
  mutate(estimate = exp(estimate),
         conf.low = exp(conf.low),
         conf.high = exp(conf.high),
         hu_char = gsub('.*:', '', term), 
         age = case_when(hu_char == "<5" ~ "< 5y",
                         hu_char == "5-15" ~ "5 - 15y"),
         hu_char = case_when(hu_char == 'Yes' ~ 'Net\nvs. no net (ref)',
                             hu_char == "positive" ~ 'Pf infection vs.\nNo Pf (ref)',
                             hu_char == "<5" ~ "Age vs.\n> 15y (ref)",
                             hu_char == "5-15" ~ "Age vs.\n> 15y (ref)",
                             hu_char == "Male" ~ "Male vs.\nFemale (ref)"),
                             hu_char = factor(hu_char, levels = c("Age vs.\n> 15y (ref)", "Male vs.\nFemale (ref)", "Net\nvs. no net (ref)", "Pf infection vs.\nNo Pf (ref)")),
         moz_char = gsub(':.*', '', term),
         moz_char = case_when(moz_char == "speciesAn. funestus" ~ "Mosquito species:\nAn. funestus vs.\nAn. gambiae SS (ref)",
                              moz_char == "multisourceYes" ~ "Multisource bloodmeal:\n Yes vs. No (ref)",
                              moz_char == "moz_head_pf_statuspositive" ~ "Sporozoites in head-thorax:\npositive vs negative (ref)",
                              moz_char == "moz_head_pf_density" ~ "Sporozoites in\nhead-thorax"),
moz_char = factor(moz_char, levels = c("Mosquito species:\nAn. funestus vs.\nAn. gambiae SS (ref)", 
"Multisource bloodmeal:\n Yes vs. No (ref)", 
"Sporozoites in head-thorax:\npositive vs negative (ref)", "Sporozoites in\nhead-thorax"))) %>%
  select(-term)

m_out_all_spz |> 
  select(moz_char, hu_char, age, estimate, conf.low, conf.high) |> 
  mutate(estimate = case_when(moz_char == 'Sporozoites in\nhead-thorax' ~ exp(log(estimate)*100),
                              TRUE ~ estimate),
         conf.low = case_when(moz_char == 'Sporozoites in\nhead-thorax' ~ exp(log(conf.low)*100),
                              TRUE ~ conf.low),
         conf.high = case_when(moz_char == 'Sporozoites in\nhead-thorax' ~ exp(log(conf.high)*100),
                              TRUE ~ conf.high),
         moz_char = case_when(moz_char == 'Sporozoites in\nhead-thorax' ~ 'Sporozoites in\nhead-thorax per 100',
                              TRUE ~ moz_char)) 
```

## Plotting probabilities

```{r}
eff_overall <- ggeffect(m_inf_all_spz, terms = c("moz_head_pf_density[0:1000, by = 1]"))

spz_overall_prob <- ggplot(eff_overall, aes(x = x, y = predicted)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.1, color = NA) +
  scale_y_continuous(limits = c(0.5,1)) +
  scale_x_continuous(limits = c(0,750))+
  labs(x = "Sporozoites per head-thorax", y = "Probability of biting\nan infected person") +
  theme_bw() +
  theme(legend.position = c(0.7,0.2), legend.box.background = element_rect(color = "black"))
```


```{r}
eff_species <- ggeffect(m_inf_all_spz, terms = c("moz_head_pf_density[0:1000, by = 1]", "species"))

spz_species_prob <- ggplot(eff_species, aes(x = x, y = predicted, color = group)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.1, color = NA) +
  scale_y_continuous(limits = c(0.5,1)) +
  scale_x_continuous(limits = c(0,750))+
  scale_color_manual(values = c('An. gambiae SS' = 'forestgreen', 'An. funestus' = 'coral'))+
  scale_fill_manual(values = c('An. gambiae SS' = 'forestgreen', 'An. funestus' = 'coral'))+
  labs(x = "Sporozoites per head-thorax", y = "Probability of biting\nan infected person", color = "Species", fill = "Species") +
  theme_bw() +
  theme(legend.position = c(0.7,0.2), legend.box.background = element_rect(color = "black"))

eff_multi <- ggeffect(m_inf_all_spz, terms = c("moz_head_pf_density[0:1000, by = 1]", "multisource"))

spz_multi_prob <- ggplot(eff_multi, aes(x = x, y = predicted, color = group)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), alpha = 0.1, color = NA) +
  scale_y_continuous(limits = c(0.5,1)) +
  scale_x_continuous(limits = c(0,750))+
  scale_color_manual(values = c('Yes' = 'cadetblue3', 'No' = 'pink3')) +
    scale_fill_manual(values = c('Yes' = 'cadetblue3', 'No' = 'pink3')) +
  labs(x = "Sporozoites per head-thorax", y = "Probability of biting\nan infected person", color = "Multisource", fill = "Multisource") +
  theme_bw() +
  theme(legend.position = c(0.8,0.2), legend.box.background = element_rect(color = "black"))
```


# Main plot

```{r, fig.width=10, fig.height=3}
(dc_plot <- ggarrange(moz_plot + theme(legend.box.spacing = unit(0, "pt")),
          ggarrange(NULL, spz_overall_prob, ncol = 1, heights = c(1, 5)),
          nrow = 1,
          widths = c(2.5, 1),
          labels = 'AUTO'))

ggsave(plot = dc_plot, filename = 'analysis/manuscript_figures/figures/discrete_choice.png',
       width = 10, height = 3)
```

```{r, fig.width=6, fig.height=3}
(spz_moz_char <- ggarrange(spz_species_prob +
                      labs(x = ""),
                    spz_multi_prob +
                      labs(x = "", y = ""),
                    labels = 'AUTO',
                    widths = c(1.05,1)) |> 
  annotate_figure(bottom = text_grob("Sporozoites per head-thorax", hjust = 0.5, vjust = -1.5)))

ggsave(plot = spz_moz_char, filename = 'analysis/manuscript_figures/figures/spz_by_moz_char.png', width = 6, height = 3)
```




